---
title: "ESM 204 HW 3"
author: "Janelle Christensen"
date: "5/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}

# ------------
# attach packages and read in csv
# ------------

library(tidyverse)
library(janitor)
library(stargazer)
library(wesanderson)

gas_df <- read_csv("Homework 3 Data.csv") %>% 
  clean_names()

```

#### 1.

```{r}

# -----------
# graph high and low and aggregate
# -----------

gas_agg_df <- gas_df %>% 
  mutate(q_agg_gallons = q_low_gallons + q_high_gallons) %>% 
  pivot_longer(q_low_gallons:q_agg_gallons,
               names_to = "consumer",
               values_to = "quantity") %>% # pivot longer to get all curved on one graph
  mutate(quantity = quantity/1000) # divided by 1000 to make graph more readable

ggplot(data = gas_agg_df, aes(x = quantity, y = price_dollars, color = consumer)) +
  geom_point() +
  theme_light() +
  scale_color_manual(values = wes_palette("Darjeeling1")) +
  labs(x = "Quantity (thousands of gallons)",
       y = "Price ($)",
       color = "Consumer")
  

```

```{r}

gas_low_df <- gas_agg_df %>% 
  filter(consumer %in% c("q_low_gallons"))

model_low <- lm(price_dollars ~ quantity, data = gas_low_df)

gas_high_df <- gas_agg_df %>% 
  filter(consumer %in% c("q_high_gallons"))

model_high <- lm(price_dollars ~ quantity, data = gas_high_df)

gas_tot_df <- gas_agg_df %>% 
  filter(consumer %in% c("q_agg_gallons"))

model_agg <- lm(price_dollars ~ quantity, data = gas_tot_df)

# solving the aggregate curve using price of 3
# times 1000 because quantity was divided by 1000 in the beginning
q_3_agg <- ((3 - 14.5998) / -.019339) * 1000

# solving the equation p = 15.8034 - 0.0273q using price of 3
# times 1000 because quantity was divided by 1000 in the beginning
q_3_high <- ((3 - 15.8034) / -0.0273) * 1000

# solving the equation p = 11.6855 -.0661q using price of 3
# times 1000 because quantity was divided by 1000 in the beginning
q_3_low <- ((3 - 11.6855) / -0.0661) * 1000

#tringle area from above the price at 3 to the aggregate 
cons_benefit_3 <- (14.5998 - 3) * q_3_agg * .5

#triangle area from price 3 at quantity
prod_benefit_3 <-  3 * q_3_agg * .5

# given cost per gallon at quantity
env_cost_local_3 <- 1.5 * q_3_agg

env_cost_world_3 <- 0.5 * q_3_agg

```

Aggregate quantity of gallons sold at 3.00/gallon is `r format(round(q_3_agg, 2), scientific=FALSE)`

Consumer benefit at 3.00/gallon is $`r format(round(cons_benefit_3, 2), scientific=FALSE)`

Producer benefit at 3.00/gallon is $`r format(round(prod_benefit_3, 2), scientific=FALSE)`

The local environmental externality cost is $`r format(round(env_cost_local_3, 2), scientific=FALSE)`

The environmental externality cost to the rest of the world is $`r format(round(env_cost_world_3, 2), scientific=FALSE)`

#### 2. 

```{r}

high_benefit_3 <- (15.8034 - 3) * q_3_high * 0.5

low_benefit_3 <- (11.6855 - 3) * q_3_low * 0.5

```

Benefit of high income consumers is approximately $`r format(round(high_benefit_3, 2), scientific=FALSE)`

Benefit of low income consumers is approximately $`r format(round(low_benefit_3, 2), scientific=FALSE)`

#### 3.

```{r}
# quantity of where the two lines are equal to each other
# times 1000 because quantity was divided by 1000 in the beginning
q_tax_agg <- ((14.5998 - 1) / (0.019339 + 0.005)) * 1000

# quantity plugged into the equation 0.005q + 1
# 0.005q was the original slope of the line, plus 1 for the $1/gallon
p_tax_agg <- 551.1811 * .005 + 1

# solving for p using p = 15.8034 - 0.0273Q
q_tax_high <- ((p_tax_agg - 15.8034) / -0.0273)

# solving for p using p = 11.8655 - 0.0661Q
q_tax_low <- ((p_tax_agg - 11.6855) / -0.0661)

# y intercept of high minus tax
#triangle area
high_benefit_tax <- (15.8034 - p_tax_agg) * q_tax_high * 0.5

# y intercept of low minus tas
# triangle area
low_benefit_tax <- (11.6855 - p_tax_agg) * q_tax_low * 0.5

# new price with tax times new quantity with tax
#triangle area
prod_benefit_tax <- p_tax_agg * q_tax_agg * .5

# same environmental cost per gallon, new quantity
env_cost_local_tax <- 1.5 * q_tax_agg

env_cost_world_tax <- 0.5 * q_tax_agg

# rectangle area from the increase in price from before to now that the tax exists

tax_rev <- 0.76 * q_tax_agg


```

a. Aggregate quantity of gallons sold at 3.00/gallon is `r format(round(q_tax_agg, 0), scientific=FALSE)`

b. Price of gasoline is $`r format(round(p_tax_agg, 2), scientific=FALSE)`

c. Benefit of high income consumers is approximately $`r format(round(high_benefit_tax, 2), scientific=FALSE)`

d. Benefit of low income consumers is approximately $`r format(round(low_benefit_tax, 2), scientific=FALSE)`

e. Producer benefit at 3.00/gallon is $`r format(round(prod_benefit_tax, 2), scientific=FALSE)`

f. The local environmental externality cost is $`r format(round(env_cost_local_tax, 2), scientific=FALSE)`

g. The environmental externality cost to the rest of the world is $`r format(round(env_cost_world_tax, 2), scientific=FALSE)`

h. Tax revenue is $`r format(round(tax_rev, 2), scientific=FALSE)`

